### 2017-07-04 reasoning.tex
Edits.

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It relies on the
transaction-local judgment to verify the the transaction body ($c$).
The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_l$) for the local judgment is a ternary relation computed as $R$
modulo $\I$ as following:
\begin{smathpar}
\begin{array}{lcl}
\R_l(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
\I\,\C{true}\,(\stl,\stg,\stg')
\end{array}
\end{smathpar}
Thus, $\R_l$ allows an interference only if it doesn't violate the
execution-time isolation specification of the transaction. If a
certain interference violates the isolation spec, i.e., $R(\stg,\stg')
\Rightarrow \neg(\I\,\C{true}\,(\stl,\stg,\stg'))$, then
$\R_l(\stl,\stg,\stg') \Leftrightarrow false$, and any assertion is
trivially stable w.r.t that interference. This is sensible considering
such interference is pre-empted in the operational semantics.

@@

For the guarantee $G$ of the transaction to be valid, it must follow
from the post-condition $Q$ of the body, provided that $Q$ is stable
w.r.t the commit-time interference captured by $R_c$. $R_c$, like
$R_l$, is computed as rely relation modulo isolation, except that
commit-time isolation ($\I\,\C{false}$) is considered. The validity of
$G$ is captured by the following implication:
\begin{smathpar}
\begin{array}{c}
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow G(\stg, \stl \gg \stg)\spc
\end{array}
\end{smathpar}
In other words, if $Q$ relates the transaction-local database state
($\stl$) to the state of the global database ($\stg$) before the
commit, then $G$ must relate the states of the global database before
and after the commit. The act of commit captured by the flush
operation ($\stl\gg\stg$). Once we establish the validity of $G$ as a
faithful representative of the transaction, we can verify that the
transaction preserves the high-level invariant $I$ by checking the
stability of $I$ w.r.t $G$, i.e., $\forall \stg,\stg'.~I(\stg) \wedge
G(\stg,\stg') \Rightarrow I(\stg')$.

@@

The \rulelabel{RG-Conseq} rule lets us safely strengthen the guarantee
$G$, or weaken the rely $R$ of a transaction. Importantly, it also
allows its isolation specification $\I$ to be strengthened. This means
that a transaction proven correct under a weaker isolation level is
also correct under a stronger level. Parametricity over the isolation
specification $\I$, combined with the ability to strengthen $\I$ as
needed, admits a flexible proof strategy to prove database programs
correct. For example, programmers can declare isolation requirements
of their choice through $\I$, and then prove programs correct assuming
the guarantees hold. The soundness of strengthening $\I$ ensures that
a program can be safely executed on any system that offers isolation
guarantees at least as strong as those assumed.

>>>>

\rulelabel{RG-Txn} is the top-level rule that lets us prove a
transaction preserves the high-level invariant $I$ when executed under
the required isolation as specified by $\I$. It relies on the
transaction-local judgment to verify the the transaction body ($c$).
The precondition $P$ of $c$ must follow from the fact that the
transaction-local database ($\stl$) is initially empty, and the global
database satisfies the high-level invariant $I$. The rely relation
($\R_l$) for the local judgment is a ternary relation computed as $R$
modulo $\I_e$ as following:
\begin{smathpar}
\begin{array}{lcl}
\R_l(\stl,\stg,\stg') & \Leftrightarrow & R(\stg,\stg') \conj
\I_e(\stl,\stg,\stg')
\end{array}
\end{smathpar}
Thus, $\R_l$ allows an interference only if it doesn't violate the
execution-time isolation specification ($\I_e$) of the transaction. If a
certain interference violates the isolation spec, i.e., $R(\stg,\stg')
\Rightarrow \neg(\I_e(\stl,\stg,\stg'))$, then $\R_l(\stl,\stg,\stg')
\Leftrightarrow false$, and any assertion is trivially stable w.r.t
that interference. This is sensible considering such interference is
pre-empted in the operational semantics.

@@

For the guarantee $G$ of the transaction to be valid, it must follow
from the post-condition $Q$ of the body, provided that $Q$ is stable
w.r.t the commit-time interference captured by $R_c$. $R_c$, like
$R_l$, is computed as rely relation modulo isolation, except that
commit-time isolation ($\I_c$) is considered. The validity of
$G$ is captured by the following implication:
\begin{smathpar}
\begin{array}{c}
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow G(\stg, \stl \gg \stg)\spc
\end{array}
\end{smathpar}
In other words, if $Q$ relates the transaction-local database state
($\stl$) to the state of the global database ($\stg$) before the
commit, then $G$ must relate the states of the global database before
and after the commit. The act of commit captured by the flush
operation ($\stl\gg\stg$). Once we establish the validity of $G$ as a
faithful representative of the transaction, we can verify that the
transaction preserves the high-level invariant $I$ by checking the
stability of $I$ w.r.t $G$, i.e., $\forall \stg,\stg'.~I(\stg) \wedge
G(\stg,\stg') \Rightarrow I(\stg')$.

@@

The \rulelabel{RG-Conseq} rule lets us safely weaken the guarantee
$G$, and strengthen the rely $R$ of a transaction. Importantly, it
also allows its isolation specification $\I$ to be strengthened (both
$\I_e$ and $\I_c$). This means that a transaction proven correct under
a weaker isolation level is also correct under a stronger level.
Parametricity over the isolation specification $\I$, combined with the
ability to strengthen $\I$ as needed, admits a flexible proof strategy
to prove database programs correct. For example, programmers can
declare isolation requirements of their choice through $\I$, and then
prove programs correct assuming the guarantees hold. The soundness of
strengthening $\I$ ensures that a program can be safely executed on
any system that offers isolation guarantees at least as strong as
those assumed.
